---
title: "Oral Microbiome - Most abundant across site `r Sys.Date()`"
author: "Florentin CONSTANCIAS"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 4
---

<style type="text/css">
div.main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
.Table-Normal {
position: relative;
//display: block;
margin: 10px auto;
padding: 0;
width: 100%;
height: auto;
border-collapse: collapse;
text-align: center;
}
</style>


```{r setup, include=FALSE}
rm(list = ls())

# knitr::opts_knit$set(root.dir = "/Users/fconstan/Projects/Oral/") 
knitr::opts_chunk$set(echo = TRUE)
DT::datatable(matrix())
# knitr::opts_chunk$set(dev = 'pdf')
# knitr::purl("Mdown.Rmd")

```


```{r packages, results=TRUE, warning=TRUE, include=FALSE}
require(tidyverse); packageVersion("tidyverse")
require(phyloseq); packageVersion("phyloseq")
```

# Fin most abundant taxa across sites:

```{r, message = FALSE, warning = FALSE, include=FALSE}
biobakery_function <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_biobakery_functions.R")
alpha <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_alpha.R")
beta <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_beta.R")
taxa <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_taxa_tests.R")
norm <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_normalisation.R")
heat <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_heatmap.R")
varia <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_varia.R")

```

```{r, message = FALSE, results= "hide", warning = FALSE, include = TRUE}
sapply(c(biobakery_function, alpha, beta, taxa, norm, heat, varia), source, chdir = TRUE)
```


# Data loading:

```{r, message = FALSE, results=FALSE}
input <- list(metpahlan_ps = "~/Documents/GitHub/oral-microbiome/data/processed/metaphlan_ps.RDS",
              motus_ps = "~/Documents/GitHub/oral-microbiome/data/processed/motus/physeq.RDS",
              pathway_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_pw_ps.RDS",
              kegg_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_kegg_ps.RDS",
              l4c_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_l4c_ps.RDS",
              go_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_go_ps.RDS",
              infogo_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_infogo_ps.RDS",
              metac_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_metac_ps.RDS",
              pfam_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_pfam_ps.RDS")

```


```{r, warning =FALSE, message = FALSE}
purrr::map(input, 
           readRDS) -> objects
```

# Top taxa based on metaphlan:

Investigate most abundant taxa over sites:

```{r, message = FALSE, warning = FALSE}
sample_data(objects$metpahlan_ps$physeq)$Site_Health <- paste0(sample_data(objects$metpahlan_ps$physeq)$Oral_Site,
                                                               "_",
                                                               sample_data(objects$metpahlan_ps$physeq)$Health_status)
```

## Overall:
```{r, message = FALSE, warning = FALSE}

physeq_most_abundant(physeq = objects$metpahlan_ps$physeq,
                     group_var = "Site_Health",
                     ntax = 3) -> most_abundant

most_abundant
```

Heatmap those:
```{r fig.height = 5 , fig.width = 12 , message=FALSE}
objects$metpahlan_ps$physeq %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  phyloseq::subset_taxa(Species %in% most_abundant) %>%
  phyloseq_ampvis_heatmap(transform = FALSE,
                          group_by = "Subject",#treatment
                          facet_by = c("Oral_Site", "Health_status", "Subject"),
                          ntax = 100,
                          tax_aggregate = "Species",
                          tax_add = NULL) -> p

p + facet_grid(. ~ Oral_Site + Health_status, scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 25,50, 75, 100), 
                       trans = scales::pseudo_log_trans(sigma = 0.1))  + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) + 
  theme(axis.text.y = element_text(angle = 0,  size = 8))
```



## Saliva & Tongue:

Test bacterial activity are compatible in Saliva vs local site only in health:

```{r message=FALSE, warnings = FALSE}
objects$metpahlan_ps$physeq %>%
  subset_samples(Health_status == "Healthy control") %>%
  subset_samples(Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) -> tmp

tmp %>%
  physeq_most_abundant(physeq = .,
                     group_var = "Oral_Site",
                     ntax = 5) -> ma_saliva_tongue

ma_saliva_tongue
```


Heatmap those:
```{r fig.height = 5 , fig.width = 12 , message=FALSE}
tmp %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  phyloseq::subset_taxa(Species %in% ma_saliva_tongue) %>%
  phyloseq_ampvis_heatmap(transform = FALSE,
                          group_by = "Subject",#treatment
                          facet_by = c("Oral_Site", "Health_status", "Subject"),
                          ntax = 100,
                          tax_aggregate = "Species",
                          tax_add = NULL) -> p

p + facet_grid(. ~ Oral_Site + Health_status, scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 25,50, 75, 100), 
                       trans = scales::pseudo_log_trans(sigma = 0.1))  + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) + 
  theme(axis.text.y = element_text(angle = 0,  size = 8))
```
Looks good. We want to check their activities among those two sites.

# Load pathway abundance table and have a look:

```{r message=FALSE, warnings = FALSE}
objects$pathway_ps$physeq  %>%
  humann2_species_contribution(meta_data_var = c("Subject", "sample_Sample" ,"Type", "Oral_Site", "Health_status")) %>%
  separate(Feature, into = c("code", "name"), sep = ": ", remove = FALSE) %>%
  dplyr::filter(!is.na(name))  -> pwy

pwy %>%
  distinct(name, .keep_all = TRUE) %>%
  DT::datatable()
```

## Visualize  RNA/DNA pathway expression:


```{r message=FALSE, warnings = FALSE}
pwy %>%
  dplyr::filter(Health_status == "Healthy control",
                Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
  humann2_RNA_DNA_plot(x_plot = "log10(DNA)", 
                       y_plot = "log10(RNA)", 
                       color = "Species", 
                       fill = "Species",
                       shape = "Genus",
                       group = "Species",
                       filter_species = ma_saliva_tongue,
                       facet_formula = "Oral_Site ~ .") -> plot


```                      

```{r, message=FALSE}
plot$legend
```
'triangle plots'


```{r fig.height = 6 , fig.width = 16 , message=FALSE}
plot$plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d()
```
```{r fig.height = 6 , fig.width = 16 , message=FALSE}
# plot$plot %>%
#   plotly::ggplotly()
```

Boxplot of each Species at each site based on overall pathway.

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
pwy %>%
  dplyr::filter(Health_status == "Healthy control",
                Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
  # filter(RNA_DNA > mean(RNA_DNA, na.rm = TRUE)) %>%
  # dplyr::filter(grepl("ose",name)) %>%
  humann2_RNA_DNA_ratio_plot(x_plot = "Species", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_species = ma_saliva_tongue,
                       facet_formula = ". ~ .",
                       export_legend = TRUE) -> plot

plot$legend
```


```{r fig.height = 10 , fig.width = 10 , message=FALSE}
plot$plot + ggpubr::rotate()

```

Test differences among species and between sites:

```{r , message=FALSE}
plot$plot$data %>% 
  ggpubr::compare_means(RNA_DNA ~ Oral_Site,
                      group.by = c("Species"),
                      data = .,
                      method = "wilcox.test",
                      p.adjust.method = "fdr") %>%
  filter(p.adj < 0.05)  %>%
  DT::datatable()
```
From here we can focus at Strep. parasanguinis and check the pathways expressed by patients in diff/ sites:


```{r,  message=FALSE}
pwy %>%
  dplyr::filter(Health_status == "Healthy control",
                Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_species = "Streptococcus_parasanguinis",
                       facet_formula = " . ~ Health_status",
                       export_legend = TRUE) -> plot
  
plot$legend
```
```{r fig.height = 10 , fig.width = 10 , message=FALSE}
plot$plot
```

Test for signicant differences and same figure
```{r , message=FALSE}
plot$plot$data %>% 
  mutate(log2RNA_DNA = log2(RNA_DNA)) %>%
  ggpubr::compare_means(log2RNA_DNA ~ Oral_Site,
                      group.by = c("Feature"),
                      data = .,
                      # method = "wilcox.test",
                      p.adjust.method = "fdr") %>%
  filter(p.adj < 0.05) -> diff_signif  

diff_signif %>%
  DT::datatable()
```


```{r , message=FALSE}
diff_signif$Feature
```

```{r fig.height = 10 , fig.width = 10 , message=FALSE}
pwy %>%
  dplyr::filter(Health_status == "Healthy control",
                Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_feature = diff_signif$Feature,
                       filter_species = "Streptococcus_parasanguinis",
                       facet_formula = " . ~ Health_status",
                       export_legend = TRUE) -> plot
  
plot$plot
```

```{r , message=FALSE}
pwy %>%
  # dplyr::filter(Health_status == "Healthy control",
   dplyr::filter(Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Health_status", 
                       fill = "Health_status",
                       shape = NULL,
                       filter_species = "Streptococcus_parasanguinis",
                       facet_formula = " . ~ Oral_Site ",
                       export_legend = TRUE) -> plot
  
plot$legend
```


```{r fig.height = 10 , fig.width = 10 , message=FALSE}
plot$plot
```

```{r , message=FALSE}
plot$plot$data %>% 
  mutate(log2RNA_DNA = log2(RNA_DNA)) %>%
  ggpubr::compare_means(log2RNA_DNA ~  Health_status,
                      group.by = c("Oral_Site","Feature"),
                      data = .,
                      # method = "wilcox.test",
                      p.adjust.method = "fdr") %>%
  filter(p.adj < 0.05) -> diff_signif  

diff_signif %>%
  DT::datatable()
```
```{r , message=FALSE}
diff_signif$Feature
```

```{r fig.height = 10 , fig.width = 10 , message=FALSE}
pwy %>%
  # dplyr::filter(Health_status == "Healthy control",
     # dplyr::filter(Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
   dplyr::filter(Oral_Site %in% c("TONGUE_BIOFILM")) %>% 
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Health_status", 
                       fill = "Health_status",
                       shape = NULL,
                       filter_species = "Streptococcus_parasanguinis",
                       filter_feature = diff_signif %>% filter(Oral_Site == "TONGUE_BIOFILM") %>% pull(Feature),
                       facet_formula = " . ~ Oral_Site ",
                       export_legend = TRUE) -> plot
  

plot$legend
plot$plot
```

```{r fig.height = 10 , fig.width = 10 , message=FALSE}
pwy %>%
  # dplyr::filter(Health_status == "Healthy control",
     # dplyr::filter(Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
   dplyr::filter(Oral_Site %in% c("SALIVA")) %>% 
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Health_status", 
                       fill = "Health_status",
                       shape = NULL,
                       filter_species = "Streptococcus_parasanguinis",
                       filter_feature = diff_signif %>% filter(Oral_Site == "SALIVA") %>% pull(Feature),
                       facet_formula = " . ~ Oral_Site ",
                       export_legend = TRUE) -> plot
  

plot$legend
plot$plot
```


# See from streptococci point of view.

First have a look at the top Genus

```{r message=FALSE, warnings = FALSE}
objects$metpahlan_ps$physeq %>%
  subset_samples(Health_status == "Healthy control") %>%
  subset_samples(Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) -> tmp

tmp %>%
  physeq_most_abundant(physeq = .,
                     group_var = "Oral_Site",
                     ntax = 3,
                     tax_level = "Genus") -> ma_saliva_tongue_g

ma_saliva_tongue_g
```


Heatmap those:
```{r fig.height = 5 , fig.width = 12 , message=FALSE}
tmp %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  phyloseq::subset_taxa(Genus %in% ma_saliva_tongue_g) %>%
  phyloseq_ampvis_heatmap(transform = FALSE,
                          group_by = "Subject",#treatment
                          facet_by = c("Oral_Site", "Health_status", "Subject"),
                          ntax = 100,
                          tax_aggregate = "Genus",
                          tax_add = NULL) -> p

p + facet_grid(. ~ Oral_Site + Health_status, scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 25,50, 75, 100), 
                       trans = scales::pseudo_log_trans(sigma = 0.1))  + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) + 
  theme(axis.text.y = element_text(angle = 0,  size = 8))
```

strep. species?


```{r message=FALSE, warnings = FALSE}
objects$metpahlan_ps$physeq %>%
  subset_samples(Health_status == "Healthy control") %>%
  subset_samples(Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) -> tmp

tmp %>%
  subset_taxa(Genus %in% c("Streptococcus")) %>%
  physeq_most_abundant(physeq = .,
                     group_var = "Oral_Site",
                     ntax = 5,
                     tax_level = "Species") -> ma_saliva_tongue_strep

ma_saliva_tongue_strep
```

```{r fig.height = 5 , fig.width = 12 , message=FALSE}
tmp %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  phyloseq::subset_taxa(Species %in% ma_saliva_tongue_strep) %>%
  phyloseq_ampvis_heatmap(transform = FALSE,
                          group_by = "Subject",#treatment
                          facet_by = c("Oral_Site", "Health_status", "Subject"),
                          ntax = 100,
                          tax_aggregate = "Species",
                          tax_add = NULL) -> p

p + facet_grid(. ~ Oral_Site + Health_status, scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 25,50, 75, 100), 
                       trans = scales::pseudo_log_trans(sigma = 0.1))  + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) + 
  theme(axis.text.y = element_text(angle = 0,  size = 8))
```

let's focus on the three most abundant:

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
strep_sel <- c("Streptococcus_infantis", "Streptococcus_parasanguinis", "Streptococcus_salivarius")
```

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
pwy %>%
  dplyr::filter(Health_status == "Healthy control",
                Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
  # filter(RNA_DNA > mean(RNA_DNA, na.rm = TRUE)) %>%
  # dplyr::filter(grepl("ose",name)) %>%
  humann2_RNA_DNA_ratio_plot(x_plot = "Species", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_species = strep_sel,
                       facet_formula = ". ~ .",
                       export_legend = TRUE) -> plot

plot$legend
```


```{r fig.height = 10 , fig.width = 10 , message=FALSE}
plot$plot + ggpubr::rotate()
```


Perform stats

```{r , message=FALSE}
plot$plot$data %>% 
  ggpubr::compare_means(RNA_DNA ~ Oral_Site,
                      group.by = c("Species"),
                      data = .,
                      method = "wilcox.test",
                      p.adjust.method = "fdr") %>%
  filter(p.adj < 0.05)  %>%
  DT::datatable()
```
Both Streptococcus_parasanguinis and Streptococcus_infantis overall display differential transicpromic activities among sites.


```{r , message=FALSE}
pwy %>%
  dplyr::filter(Health_status == "Healthy control",
                Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_species = strep_sel,
                       facet_formula = " . ~ Health_status + Species",
                       export_legend = TRUE) -> plot
  
plot$legend
```

```{r fig.height = 12 , fig.width = 18 , message=FALSE}
plot$plot
```

We could pintpoint pathways of interests here.
-test for significant diff + health vs perio at heach sites
-stay at pathway level before going deeper.

```{r , message=FALSE}
plot$plot$data %>% 
  # mutate(log2RNA_DNA = log2(RNA_DNA)) %>%
  ggpubr::compare_means(RNA_DNA ~  Oral_Site,
                      group.by = c("Species","Feature"),
                      data = .,
                      # method = "wilcox.test",
                      p.adjust.method = "fdr") %>%
  filter(p.adj < 0.05) -> diff_signif  

diff_signif %>%
  DT::datatable()
```

```{r fig.height = 10 , fig.width = 10 , message=FALSE}
pwy %>%
  dplyr::filter(Health_status == "Healthy control") %>%
     dplyr::filter(Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
   # dplyr::filter(Oral_Site %in% c("TONGUE_BIOFILM")) %>% 
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_species = "Streptococcus_infantis",
                       filter_feature = diff_signif %>% filter(Species == "Streptococcus_infantis") %>% pull(Feature),
                       facet_formula = " . ~ Health_status ",
                       export_legend = TRUE) -> plot
  

plot$legend
plot$plot
```
```{r fig.height = 10 , fig.width = 10 , message=FALSE}
pwy %>%
  dplyr::filter(Health_status == "Healthy control") %>%
     dplyr::filter(Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
   # dplyr::filter(Oral_Site %in% c("TONGUE_BIOFILM")) %>% 
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_species = "Streptococcus_parasanguinis",
                       filter_feature = diff_signif %>% filter(Species == "Streptococcus_parasanguinis") %>% pull(Feature),
                       facet_formula = " . ~ Health_status ",
                       export_legend = TRUE) -> plot
  

plot$legend
plot$plot
```
Streptococcus_salivarius

```{r fig.height = 10 , fig.width = 10 , message=FALSE}
pwy %>%
  dplyr::filter(Health_status == "Healthy control") %>%
     dplyr::filter(Oral_Site %in% c("SALIVA", "TONGUE_BIOFILM")) %>%
   # dplyr::filter(Oral_Site %in% c("TONGUE_BIOFILM")) %>% 
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_species = "Streptococcus_salivarius",
                       filter_feature = diff_signif %>% filter(Species == "Streptococcus_salivarius") %>% pull(Feature),
                       facet_formula = " . ~ Health_status ",
                       export_legend = TRUE) -> plot
  

plot$legend
plot$plot
```

