---
title: "Oral Microbiome - Most abundant across site `r Sys.Date()`"
author: "Florentin CONSTANCIAS"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 4
---

<style type="text/css">
div.main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
.Table-Normal {
position: relative;
//display: block;
margin: 10px auto;
padding: 0;
width: 100%;
height: auto;
border-collapse: collapse;
text-align: center;
}
</style>


```{r setup, include=FALSE}
rm(list = ls())

# knitr::opts_knit$set(root.dir = "/Users/fconstan/Projects/Oral/") 
knitr::opts_chunk$set(echo = TRUE)
DT::datatable(matrix())
# knitr::opts_chunk$set(dev = 'pdf')
# knitr::purl("Mdown.Rmd")

```


```{r packages, results=TRUE, warning=TRUE, include=FALSE}
require(tidyverse); packageVersion("tidyverse")
require(phyloseq); packageVersion("phyloseq")
```

# Fin most abundant taxa across sites:

```{r, message = FALSE, warning = FALSE}
biobakery_function <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_biobakery_functions.R")
alpha <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_alpha.R")
beta <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_beta.R")
taxa <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_taxa_tests.R")
norm <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_normalisation.R")
heat <- ("~/Documents/GitHub/DivComAnalyses/R/phyloseq_heatmap.R")
```

```{r, message = FALSE, results= "hide", warning = FALSE, include = TRUE}
sapply(c(biobakery_function, alpha, beta, taxa, norm, heat), source, chdir = TRUE)
```

```{r, message = FALSE, results=FALSE}
input <- list(metpahlan_ps = "~/Documents/GitHub/oral-microbiome/data/processed/metaphlan_ps.RDS",
              motus_ps = "~/Documents/GitHub/oral-microbiome/data/processed/motus/physeq.RDS",
              pathway_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_pw_ps.RDS",
              kegg_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_kegg_ps.RDS",
              l4c_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_l4c_ps.RDS",
              go_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_go_ps.RDS",
              infogo_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_infogo_ps.RDS",
              metac_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_metac_ps.RDS",
              pfam_ps = "~/Documents/GitHub/oral-microbiome/data/processed/humann_pfam_ps.RDS")

```

Data loading:

```{r, warning =FALSE, message = FALSE}
purrr::map(input, 
           readRDS) -> objects
```

# metaphlan Taxa:

Investigate most abundant taxa over sites:

```{r, message = FALSE, warning = FALSE}
sample_data(objects$metpahlan_ps$physeq)$Site_Health <- paste0(sample_data(objects$metpahlan_ps$physeq)$Oral_Site,
                                                               "_",
                                                               sample_data(objects$metpahlan_ps$physeq)$Health_status)
```

```{r, message = FALSE, warning = FALSE}

physeq_most_abundant <- function(physeq,
                                 group_var,
                                 ntax = 10,
                                 tax_level = "Species"){
  
  require(tidyverse); require(phyloseq)
  
  taxa_top_all = NULL  
  
  for(tp in physeq %>% 
      get_variable(group_var) %>%
      unique()){
    # print(tp)
    prune_samples(get_variable(physeq, group_var) == tp,
                  physeq) %>%
      fantaxtic::get_top_taxa(n = ntax, 
                              relative = FALSE, 
                              discard_other = TRUE) -> tmp2
    
    as(tax_table(tmp2), "matrix") %>%
      data.frame() %>%
      # dplyr::filter(grepl("ose",Gene_name)) %>%
      pull(tax_level) -> spc
    
    c(spc, taxa_top_all) %>%
      # discard(is.na)  %>%
      unique() %>%
      sort() -> taxa_top_all
    
  }
  return(taxa_top_all)
}


physeq_most_abundant(physeq = objects$metpahlan_ps$physeq,
                     group_var = "Site_Health",
                     ntax = 3) -> most_abundant

most_abundant
```

Heatmap those:
```{r fig.height = 5 , fig.width = 12 , message=FALSE}
objects$metpahlan_ps$physeq %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  phyloseq::subset_taxa(Species %in% most_abundant) %>%
  phyloseq_ampvis_heatmap(transform = FALSE,
                          group_by = "Subject",#treatment
                          facet_by = c("Oral_Site", "Health_status", "Subject"),
                          ntax = 100,
                          tax_aggregate = "Species",
                          tax_add = NULL) -> p

p + facet_grid(. ~ Oral_Site + Health_status, scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 25,50, 75, 100), 
                       trans = scales::pseudo_log_trans(sigma = 0.1))  + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) + 
  theme(axis.text.y = element_text(angle = 0,  size = 8))
```

What are those species doing:

# PAthway

```{r message=FALSE, warnings = FALSE}
objects$pathway_ps$physeq  %>%
  humann2_species_contribution(meta_data_var = c("Subject", "sample_Sample" ,"Type", "Oral_Site", "Health_status")) %>%
  separate(Feature, into = c("code", "name"), sep = ": ", remove = FALSE) %>%
  dplyr::filter(!is.na(name))  -> tmp

tmp %>%
  distinct(name, .keep_all = TRUE) %>%
  DT::datatable()
```

Distribution

```{r message=FALSE, warnings = FALSE}
tmp %>%
  dplyr::select(-id) %>%
  # dplyr::filter(is.numeric(DNA) & is.numeric(RNA) & is.numeric(RNA_DNA) &  RNA_DNA != Inf) %>%
  # dplyr::filter(DNA > 0, DNA > 0, abs(RNA_DNA) > 0) %>%
  pivot_longer(cols = c("DNA", "RNA", "RNA_DNA"),
               values_to = "value",
               names_to = 'Type') %>%
  ggplot(aes(y=value, x=Type, color = Type, fill = Type)) +
  geom_boxplot(outlier.colour = NA, alpha = 0.2,
               position = position_dodge(width=0.7)) +
  # geom_jitter(size=1, alpha=0.2) +
  # ggbeeswarm::geom_beeswarm(size=1, alpha=0.2) +
  scale_y_log10() + ylab("log10(prop)")

```

```{r message=FALSE, warnings = FALSE}
# objects$pathway_ps$physeq %>%
#   phyloseq_get_humann_strat_un_output(output = "stratified", # stratified / unstratified
#                                       remove_unmapped_unintegrated = FALSE, 
#                                       transform = "compositional",
#                                       export_long_df = TRUE) -> out$df
# 
# 
# out$df

```

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
tmp %>%
  group_by(Species, Subject) %>%   
  top_n(5, wt = RNA_DNA) %>%
  ungroup() %>%
  # filter(RNA_DNA > mean(RNA_DNA, na.rm = TRUE)) %>%
  # dplyr::filter(grepl("ose",name)) %>%
  humann2_RNA_DNA_ratio_plot(x_plot = "name", 
                       y_plot = "log2(RNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_species = most_abundant,
                       facet_formula = " . ~ Health_status") -> plot
```

```{r fig.height = 60 , fig.width = 10 , message=FALSE, warning= FALSE}
# plot$legend

plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d()
```

```{r message=FALSE, warnings = FALSE}
tmp %>%
  humann2_RNA_DNA_plot(x_plot = "log2(DNA)", 
                       y_plot = "log2(RNA)", 
                       color = "Feature", 
                       fill = "Feature",
                       shape = "Genus",
                       group = "Feature",
                       filter_species = most_abundant,
                       facet_formula = "Oral_Site ~ Health_status + Subject") -> plot
```

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
# plot$legend

plot$plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d()
```


```{r fig.height = 30, fig.width = 30 , message=FALSE}
# plot$legend

plot$plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d() + 
  facet_null() + facet_grid(as.formula("Oral_Site + Species ~ Health_status + Subject"), 
                   scale="fixed",space="free_x")
```

```{r fig.height = 30, fig.width = 30 , message=FALSE}
# plot$legend

plot$plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d() + 
  facet_null() + facet_grid(as.formula("Oral_Site + Species ~ Health_status + Subject"), 
                   scale="fixed",space="free_x")
```

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
# plot$legend

plot$plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d() + 
  facet_null() + 
  facet_grid(as.formula("Oral_Site ~ Health_status + Species"), drop=TRUE,
                   scale="fixed",space="free_x") 
```


```{r fig.height = 6 , fig.width = 16 , message=FALSE}
tmp %>%
  dplyr::filter(grepl("fermentation",name)) %>%
  humann2_RNA_DNA_plot(x_plot = "log2(DNA)", 
                       y_plot = "log2(RNA)", 
                       color = "Health_status", 
                       fill = "Health_status",
                       shape = "Oral_Site",
                       group = "Subject",
                       filter_species = most_abundant,
                       facet_formula = " Species ~ name") -> plot
```

```{r fig.height = 6 , fig.width = 6 , message=FALSE}
# plot$legend

plot$plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d()
```

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
tmp %>%
  dplyr::filter(grepl("ose",name)) %>%
  humann2_RNA_DNA_plot(x_plot = "log2(DNA)", 
                       y_plot = "log2(RNA)", 
                       color = "Health_status", 
                       fill = "Health_status",
                       shape = "Oral_Site",
                       group = "Subject",
                       filter_species = most_abundant,
                       facet_formula = " Species ~ name") -> plot
```

```{r fig.height = 6 , fig.width = 6 , message=FALSE}
# plot$legend

plot$plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d()
```

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
tmp %>%
  dplyr::filter(grepl("ose",name)) %>%
  humann2_RNA_DNA_ratio_plot(x_plot = "Health_status", 
                       y_plot = "log2(RNA_DNA)", 
                       color = "Oral_Site", 
                       fill = "Oral_Site",
                       shape = NULL,
                       filter_species = most_abundant,
                       facet_formula = " Species ~ name") -> plot
```

```{r fig.height = 20 , fig.width = 20 , message=FALSE}
# plot$legend

plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d()
```

```{r message=FALSE, warnings = FALSE}
tmp %>%
  dplyr::filter(RNA > DNA) %>%
  humann2_RNA_DNA_plot(x_plot = "log2(DNA)", 
                       y_plot = "log2(RNA)", 
                       color = "Species", 
                       fill = "Species",
                       shape = "Genus",
                       group = "Subject",
                       filter_species = most_abundant,
                       facet_formula = "Oral_Site  ~ Health_status + Subject") -> plot
```

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
plot$legend

plot$plot +
  scale_fill_viridis_d() + 
  scale_color_viridis_d()
```

```{r fig.height = 30 , fig.width = 12 , message=FALSE}

as.list(setNames(most_abundant, most_abundant)) -> list

lapply(
  list,
  FUN = humann2_RNA_DNA_ratio_plot_lapply,
  df = tmp,
  x_plot = "Feature",
  y_plot = "log2(RNA_DNA)",
  color = "Health_status",
  fill = "Health_status",
  filter_feature = FALSE,
  export_legend = FALSE,
  facet_formula = ". ~ Oral_Site ") -> out
```

```{r fig.height = 12 , fig.width = 8 , message=FALSE}
for (tax in c(names(list)))
{
  out[[tax]] + scale_x_discrete(label = function(x) stringr::str_trunc(x, 50, 
                                                                       side = "center")) -> p
  p %>%
    plot()
}
```

```{r fig.height = 30 , fig.width = 12 , message=FALSE}

as.list(setNames(most_abundant, most_abundant)) -> list

lapply(
  list,
  FUN = humann2_RNA_DNA_ratio_plot_lapply,
  df = tmp,
  x_plot = "Feature",
  y_plot = "log2(RNA_DNA)",
  color = "Oral_Site",
  fill = "Oral_Site",
  filter_feature = FALSE,
  export_legend = FALSE,
  facet_formula = ". ~ Health_status ") -> out

```

```{r fig.height = 12 , fig.width = 8 , message=FALSE}
for (tax in c(names(list)))
{
  out[[tax]] + scale_x_discrete(label = function(x) stringr::str_trunc(x, 50, side = "center")) -> p
  p %>%
    plot()
}
```

# metacyc

```{r message=FALSE, warnings = FALSE}
objects$metac_ps$physeq  %>%
  humann2_species_contribution(meta_data_var = c("Subject", "sample_Sample" ,"Type", "Oral_Site", "Health_status")) %>%
  separate(Feature, into = c("code", "name"), sep = ": ", remove = FALSE) %>%
  dplyr::filter(name != "NO_NAME")  -> tmp

tmp %>%
  distinct(name) %>%
  DT::datatable()
```

```{r message=FALSE, warnings = FALSE}
tmp %>%
  dplyr::select(-id) %>%
  dplyr::filter(is.numeric(DNA) & is.numeric(RNA) & is.numeric(RNA_DNA) &  RNA_DNA != Inf) %>%
  # dplyr::filter(DNA > 10, abs(RNA_DNA) > 5) %>%
  pivot_longer(cols = c("DNA", "RNA", "RNA_DNA"),
               values_to = "value",
               names_to = 'Type') %>%
  ggplot(aes(y=value, x=Type, color = Type, fill = Type)) +
  geom_boxplot(outlier.colour = NA, alpha = 0.2,
               position = position_dodge(width=0.7)) +
  # geom_jitter(size=1, alpha=0.2) +
  # ggbeeswarm::geom_beeswarm(size=1, alpha=0.2) +
  scale_y_log10() + ylab("log10(CPM)")

```
```{r fig.height = 8 , fig.width = 20 , message=FALSE}
tmp %>%
  humann2_RNA_DNA_plot(x_plot = "log2(DNA)", 
                       y_plot = "log2(RNA)", 
                       color = "Species", 
                       fill = "Species",
                       shape = "Genus",
                       group = "Species",
                       filter_species = most_abundant,
                       facet_formula = "Oral_Site ~ Health_status + Subject ") -> plot

plot
```
```{r fig.height = 10 , fig.width = 8 , message=FALSE}

as.list(setNames(most_abundant, most_abundant)) -> list

lapply(
  list,
  FUN = humann2_RNA_DNA_ratio_plot_lapply,
  df = tmp,
  x_plot = "name",
  y_plot = "log2(RNA_DNA)",
  color = "Health_status",
  fill = "Health_status",
  filter_feature = FALSE,
  export_legend = FALSE,
  facet_formula = ". ~ Oral_Site ") -> out

# out$Actinomyces_oris + 
#   scale_x_discrete(label = function(x) stringr::str_trunc(x, 50, 
#                                                                        side = "center"))
```

```{r fig.height = 20 , fig.width = 8 , message=FALSE}
for (tax in c(names(list)))
{
  out[[tax]] + scale_x_discrete(label = function(x) stringr::str_trunc(x, 40, 
                                                                       side = "center")) -> p
  p %>%
    plot()
}
```

# kegg

```{r message=FALSE, warnings = FALSE}
objects$kegg_ps$physeq  %>%
  humann2_species_contribution(meta_data_var = c("Subject", "sample_Sample" ,"Type", "Oral_Site", "Health_status"))  %>%
  separate(Feature, into = c("code", "name"), sep = ": ", remove = FALSE) %>%
  dplyr::filter(name != "NO_NAME")  -> tmp

tmp %>%
  distinct(name, .keep_all = TRUE) %>%
  DT::datatable()
```


```{r message=FALSE, warnings = FALSE}
tmp %>%
  dplyr::select(-id) %>%
  dplyr::filter(is.numeric(DNA) & is.numeric(RNA) & is.numeric(RNA_DNA) &  RNA_DNA != Inf) %>%
  # dplyr::filter(DNA > 10, abs(RNA_DNA) > 5) %>%
  pivot_longer(cols = c("DNA", "RNA", "RNA_DNA"),
               values_to = "value",
               names_to = 'Type') %>%
  ggplot(aes(y=value, x=Type, color = Type, fill = Type)) +
  geom_boxplot(outlier.colour = NA, alpha = 0.2,
               position = position_dodge(width=0.7)) +
  # geom_jitter(size=1, alpha=0.2) +
  # ggbeeswarm::geom_beeswarm(size=1, alpha=0.2) +
  scale_y_log10() + ylab("log10(CPM)")

```


```{r message=FALSE, warnings = FALSE}
tmp %>%
  humann2_RNA_DNA_plot(x_plot = "log2(DNA)", 
                       y_plot = "log2(RNA)", 
                       color = "name", 
                       fill = "name",
                       shape = "Genus",
                       group = "name",
                       filter_species = most_abundant,
                       facet_formula = "Oral_Site ~ Health_status ") -> plot
```

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
plot$plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d()
```


```{r message=FALSE, warnings = FALSE}
tmp %>%
  humann2_RNA_DNA_plot(x_plot = "log2(DNA)", 
                       y_plot = "log2(RNA)", 
                       color = "name", 
                       fill = "name",
                       shape = "Genus",
                       group = "name",
                       filter_feature = most_abundant,
                       facet_formula = "Oral_Site ~ Health_status ") -> plot
```


```{r fig.height = 30 , fig.width = 12 , message=FALSE}

as.list(setNames(most_abundant, most_abundant)) -> list

lapply(
  list,
  FUN = humann2_RNA_DNA_ratio_plot_lapply,
  df = tmp,
  x_plot = "Feature",
  y_plot = "log2(RNA_DNA)",
  color = "Health_status",
  fill = "Health_status",
  filter_feature = FALSE,
  export_legend = FALSE,
  facet_formula = ". ~ Oral_Site ") -> out
```

```{r fig.height = 12 , fig.width = 8 , message=FALSE}
for (tax in c(names(list)))
{
  out[[tax]] + scale_x_discrete(label = function(x) stringr::str_trunc(x, 50, 
                                                                       side = "center")) -> p
  p %>%
    plot()
}
```



# l4c_ps

```{r message=FALSE, warnings = FALSE}
objects$l4c_ps$physeq_count %>%
  humann2_species_contribution(meta_data_var = c("Subject", "sample_Sample" ,"Type", "Oral_Site", "Health_status"))  %>%
  separate(Feature, into = c("code", "name"), sep = ": ", remove = FALSE) %>%
  dplyr::filter(name != "NO_NAME")  -> tmp

tmp %>%
  select(-id) %>%
  distinct(name, .keep_all = TRUE) %>%
  DT::datatable()
```


```{r message=FALSE, warnings = FALSE}
tmp %>%
  dplyr::select(-id) %>%
  dplyr::filter(is.numeric(DNA) | is.numeric(RNA)) %>%
  # dplyr::filter(DNA > 10, abs(RNA_DNA) > 5) %>%
  pivot_longer(cols = c("DNA", "RNA", "RNA_DNA"),
               values_to = "value",
               names_to = 'Type') %>%
  ggplot(aes(y=value, x=Type, color = Type, fill = Type)) +
  geom_boxplot(outlier.colour = NA, alpha = 0.2,
               position = position_dodge(width=0.7)) +
  # geom_jitter(size=1, alpha=0.2) +
  # ggbeeswarm::geom_beeswarm(size=1, alpha=0.2) +
  scale_y_log10() + ylab("log10(CPM)")

```

```{r message=FALSE, warnings = FALSE}
tmp %>%
  select(-id) %>% 
  DT::datatable()
```

```{r message=FALSE, warnings = FALSE}

humann2_RNA_DNA_plot(x_plot = "log2(DNA)", 
                       y_plot = "log2(RNA)", 
                       color = "name", 
                       fill = "name",
                       shape = "Genus",
                       group = "name",
                       filter_species = most_abundant,
                       facet_formula = "Oral_Site ~ Health_status ") -> plot
```

```{r fig.height = 6 , fig.width = 16 , message=FALSE}
plot$plot  + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d()
```

```{r fig.height = 30 , fig.width = 12 , message=FALSE}

as.list(setNames(most_abundant, most_abundant)) -> list

lapply(
  list,
  FUN = humann2_RNA_DNA_ratio_plot_lapply,
  df = tmp,
  x_plot = "Feature",
  y_plot = "log2(RNA_DNA)",
  color = "Health_status",
  fill = "Health_status",
  filter_feature = FALSE,
  export_legend = FALSE,
  facet_formula = ". ~ Oral_Site ") -> out
```

```{r fig.height = 12 , fig.width = 8 , message=FALSE}
for (tax in c(names(list)))
{
  out[[tax]] + scale_x_discrete(label = function(x) stringr::str_trunc(x, 50, 
                                                                       side = "center")) -> p
  p %>%
    plot()
}
```
